---
title: "Lab02"
author: "Zaid Taiyab"
date: "2024-09-29"
output: pdf_document
---

```{r setup, include=FALSE}
options(error = recover)  # Enable error recovery
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(ggplot2movies)
library(tidyr)
data(movies)

```

# Introduction

In this lab, I will explore data about movies collected from http://imdb.com/.

## 1. What is the range of years of production of the movies?
```{r}
oldest <- min(movies$year, na.rm = TRUE)
newest <- max(movies$year, na.rm = TRUE)
range <- newest - oldest

print(paste("Oldest year:", oldest))
print(paste("Newest year:", newest))
print(paste("Range of years:", range))
```

## 2. Proportion of Movies with and without Budget, and Top 5 Most Expensive Movies
```{r}
total_movies_count <- nrow(movies)
movies_with_budget <- sum(!is.na(movies$budget))
movies_without_budget <- sum(is.na(movies$budget))
proportion_with <- movies_with_budget / total_movies_count
proportion_without <- movies_without_budget / total_movies_count

print(paste("Total number of movies:", total_movies_count))
print(paste("Movies with budget:", movies_with_budget))
print(paste("Movies without budget:", movies_without_budget))
print(paste("Proportion with budget:", round(proportion_with, 2)))
print(paste("Proportion without budget:", round(proportion_without, 2)))

top_5_expensive_movies <- movies %>%
  filter(!is.na(budget)) %>%
  arrange(desc(budget)) %>%
  head(5)

print(top_5_expensive_movies[, c("title", "budget")])
```

## 3. Top 5 Longest Movies
```{r}
longestMovies <- movies %>%
  arrange(desc(length)) %>%
  head(5)
print(longestMovies[, c("title", "length")])
```

## 4. Shortest and Longest Short Movies
```{r}
shortestMovies <- movies %>% filter(Short == 1)
if (nrow(shortestMovies) > 0) {
  valid_lengths <- shortestMovies %>% filter(!is.na(length))
  if (nrow(valid_lengths) > 0) {
    min_length <- min(valid_lengths$length)
    max_length <- max(valid_lengths$length)
    shortestMovies <- valid_lengths %>% filter(length == min_length)
    longestMovies <- valid_lengths %>% filter(length == max_length)
    print("Shortest short movie:")
    print(shortestMovies[, c("title", "length")])
    print("Longest short movie:")
    print(longestMovies[, c("title", "length")])
  } else {
    print("No short movies with valid lengths found.")
  }
} else {
  print("No valid short movies found.")
}

```

## 5. Number of Movies per Genre (Bar Plot)
```{r}

genre_counts <- colSums(movies[, c("Action", "Animation", "Comedy", 
                                   "Drama", "Documentary", "Romance", "Short")], na.rm=TRUE)
genre_data <- data.frame(genre = names(genre_counts), count = genre_counts)
ggplot(genre_data, aes(x = genre, y = count)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Movies per Genre", x = "Genre", y = "Count")

```

## 6. Average Rating per Genre (Bar Plot)
```{r}
average_ratings <- movies %>%
  pivot_longer(cols = Action:Short, names_to = "genre", values_to = "value") %>%
  
  filter(value == 1) %>%
  group_by(genre) %>%
  summarise(avg_rating = mean(rating, na.rm = TRUE))

ggplot(average_ratings, aes(x = genre, y = avg_rating)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Rating per Genre", x = "Genre", y = "Average Rating")

```

## 7. Average Rating of Movies per Genre (2000-2005)
```{r}
moviesBetween00_05 <- movies %>%
  filter(year >= 2000 & year <= 2005)

average_ratings_2000_2005 <- moviesBetween00_05 %>%
  pivot_longer(cols = Action:Short, names_to = "genre", values_to = "value") %>%
  filter(value == 1) %>%
  group_by(genre) %>%
  summarise(avg_rating = mean(rating, na.rm = TRUE), .groups = 'drop')
print(average_ratings_2000_2005)

```

## 8. Movies per Genre by Year (1990 Onwards)
```{r}
movies1990 <- movies %>% 
  filter(year >= 1990) 
genre_year_counts <- movies1990 %>%
  pivot_longer(cols = Action:Short, 
               names_to = "genre", 
               values_to = "value") %>%
  filter(value == 1, genre != "Short") %>%
  group_by(year, genre) %>%
  summarise(count = n(), .groups = 'drop')
ggplot(genre_year_counts, aes(x = year, y = count, color = genre)) +
  geom_line() +
  labs(title = "Movies per Genre by Year (1990 Onwards)", 
       x = "Year", 
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "bottom") 
```

## Question 9.1: What is the average and median rating of drama movies?

```{r}
drama_movies <- movies %>% filter(Drama == 1)
average_rating <- mean(drama_movies$rating, na.rm = TRUE)
median_rating <- median(drama_movies$rating, na.rm = TRUE)
print(paste("Average rating of drama movies:", round(average_rating, 2)))
print(paste("Median rating of drama movies:", round(median_rating, 2)))
ggplot(drama_movies, aes(x = rating)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Ratings for Drama Movies",
       x = "IMDB Rating",
       y = "Count")
```

## Question 9.2: What is the average budget of drama movies?

```{r}
drama_movies_budget <- drama_movies %>% filter(!is.na(budget))
average_budget <- mean(drama_movies_budget$budget, na.rm = TRUE)
print(paste("Average budget of drama movies:", round(average_budget, 2)))
ggplot(drama_movies_budget, aes(x = budget)) +
  geom_histogram(binwidth = 1e7, fill = "lightgreen", color = "black") +
  labs(title = "Distribution of Budgets for Drama Movies",
       x = "Budget (USD)",
       y = "Count")
```

## Question 9.3: Compare ratings of drama movies vs. other genres
```{r}
genre_ratings <- movies %>%
  pivot_longer(cols = c("Action", "Animation", "Comedy", "Drama", "Documentary", "Romance"),
               names_to = "genre", values_to = "value") %>%
  filter(value == 1) %>%
  select(title, genre, rating) %>%
  filter(!is.na(rating))
ggplot(genre_ratings, aes(x = genre, y = rating, fill = genre)) +
  geom_boxplot() +
  labs(title = "Comparison of Movie Ratings Across Genres",
       x = "Genre",
       y = "IMDB Rating") +
  theme(legend.position = "none")
```